Representing Scenes as Light Fields for Real Time Rendering in Virtual Reality Display

### 2021-01-28 TOG/4-method.tex
last day

\section{Method}
\label{sec:method}
As visualized in \Cref{fig:pipeline}, our method consists of three major components: foveal-periphery image synthesis (\Cref{sec:method:net}), and real-time blending with optimized spatial-temporal precision (\Cref{sec:method:blending}).

%\subsection{Pipeline}
\begin{figure*}[htb]
    \centering
    \includegraphics[width=\linewidth]{TOG/figs/sys_pipeline.png}
    \Caption{System pipeline.}
    {%
    }
    \label{fig:pipeline}
\end{figure*}

%To meet the requirement of real-time, use fewer channels/layers, much fewer samples, predict a coarse fovea image, then refine using local view guidance

\subsection{GazeNet}
\label{sec:method:net}
\begin{figure*}[htb]
    \centering
    \includegraphics[width=\linewidth]{TOG/figs/spher_net_flow.png}
    \caption{GazeNet Network.}
    \label{fig:my_label}
\end{figure*}

%Follow NeRF's idea, represent scene using a MLP but in spherical space because we are inside-out (viewer is located at center area and look around)

\qisun{(Jan 9, 2021) Avoid usage of subsubsection.}
\paragraph{Multi spherical representation}
As reviewed in \autoref{sec:prior}, prior work contributed a lot to represent a static scene via object-focused manner. Inspired by that, we investigate on how to represent a static scene for VR environment to render real-timely without loading the entire scene. However, VR environment is user-centered. That is, unlike the object-focused ``outside-in'' way such as \cite{mildenhall2020nerf}, the views are highly varied in an ``inside-out'' fashion. 
Consequently, the overlaps among views are diversified, resulting in training challenges for large virtual scenes.
To address this problem, we depict a VR scene as concentric spherical volumes with the center as the original point, as visualized in \Cref{fig:teaser:scene}.

%\begin{figure}[htb]
%    \centering
%    \includegraphics[width=0.96\linewidth]{TOG/figs/spher_coord.pdf}
%%    \caption{Visualization of our multi-spherical coordinate system.}
    \label{fig:method:coordinate}
%\end{figure}
% different position (10cm working, 50cm not working),
% different rotation

\dnc{(Jan 11) Ray to spherical coord} For a sphere with radius $r$, the intersection of ray and sphere $\pt_r$ can be calculated by solving:

\begin{equation}
    \|\pt_r\|^2=\|\rayo + k\rayd\|^2=r^2
\end{equation}

Convert $\pt_r$ to spherical coordinate $(r, \theta, \phi)$:

\begin{equation}
    \begin{cases}
        \theta = \mathrm{atan2}(x_{\pt_r}/z_{\pt_r})\\
        \phi = \mathrm{acos}(y_{\pt_r} / r)
    \end{cases}
\end{equation}

\dnc{(Jan 11) A simple derivation for the requirement of depth range and samples} Take 2D as an example. Consider two view centers $\rayo_1$, $\rayo_2$ on X axis and a point $\pt$ in 3D space. The radius of nearest sampling sphere of $\pt$ is $r_l$. The intersections of the sphere and rays from $\rayo_1$, $\rayo_2$ to $\pt$ are $\pt'_{1,r_l}$ and $\pt'_{2,r_l}$, respectively. $\Delta\pt'_{r_l}=\|\pt'_{2,r_l}-\pt'_{1,r_l}\|$ is small enough, then we have:

\begin{equation}
    \Delta\pt'_{r_l} \approx \Delta\rayo \frac{|z_\pt - r_l|}{z_\pt}
\end{equation}

\begin{equation}
    hello
\end{equation}

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.96\linewidth]{TOG/figs/depth_disparity.pdf}
    \caption{Depth disparity}
    \label{fig:method:eccentricity}
\end{figure}

With this coordinate system, the input of individual element in the 3D spaces is defined as $p=(r, \theta, \phi)$ and the output is color information $c=(r,g,b)$ and volume density $d$. We train the scene representation with an multilayer perception (MLP) network first. 
For the training uniformity and run-time efficiency, we do not distinguish different eccentricities during the training stage. Instead, the spatially variant visual acuity is represented as varying field of views in the rendering, i.e., data creation.
Specifically, we depict the whole visual field as three semantic layers the fovea ({0-10 deg}), mid-periphery {6-60 deg}, and far periphery {y-110 deg}. In each frame, The uniformized networks returns three images with a identical resolution yet, as seen from the separation range in deg, gradually larger areas along eccentricity.

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.96\linewidth]{TOG/figs/layers_together_zh.pdf}% try not overlapping the text on the image since we want the reviewers to see each pixel clearly
    \caption{Visualization of the output at three eccentricity ranges.}
    \label{fig:method:eccentricity}
\end{figure}

%estimate the fovea, mid-peripheral, peripheral field of view for a novel view next, fine-tune the fovea result with local view streamed from the cloud, and lastly warp/blend the three results for VR display.

\paragraph{Input encoding}
Since the input of our representation network is only three dimensional, we map the inputs to a higher dimensional space via high-frequency functions. \qisun{(Jan 9, 2021) What is bad if we have low-dim input? Harder to converge, higher error etc.?} \zh{(Jan 9) similar to NeRF, three components are not sufficient for achieving state-of-the-art quality, as demonstrated in Section 6.4). We introduce a positional encoding of the input coordinates that assists the MLP in
representing high-frequency functions.}
Thus we encoded the input through function:
\begin{align}
\iota (p) = (\sin(2^0 p), \cos(2^0 p), ..., \sin(2^{L-1}p, \cos(2^{L-1}p))
\label{eqn:inputencode}
\end{align}
The function $\iota (\cdot)$ is applied to each of the 3D location representation ($r$, $\theta$, and $\phi$), we choose $L$ to be 10 in practice. Hence, the input dimension becomes 3 x 20 = 60 after encoding. The MLP network is composed with $N_{layer}$ fully-connected layers, each uses ReLu as activation function and $N_{channel}$ channels per layer and output the color and volume density.

\paragraph{Synthesis}
% the description of the volume density is the same as chapter 4 in NeRF paper.
Similar to \cite{mildenhall2020nerf}, the trained representation is then deployed to synethesis final images via ray marching. For accelerating the performance, we bounded the near/far range to $20$ and $50$ meters respectively. The bounding, however, may introduce subtle quality drop, as shown in \Cref{fig:method:bounding}.
\begin{figure}
    \centering
    \subfloat[w bounding]{\includegraphics[width = 0.47\linewidth]{TOG/figs/depth_without_bound.png}}\hspace{1em}
    \subfloat[w/o bounding]{\includegraphics[width = 0.47\linewidth]{TOG/figs/depth_with_bound.png}}
    \caption{Visualization of ray marching bounding.}
    \label{fig:method:bounding}
\end{figure}
%For each pixel, we perform a ray from the camera and accumulate all the intersection's color weighted by the density accordingly between near bound and far bound. In practice, we experimented with different far bound values: 20 (meters) and 50 (meters), the difference between output is minimal. (\note{add figure for comparison}). We also experimented with the sampling number ($N_{sample}$) for accumulation between near bound and far bound to balance time cost and effects. Hence, we have an output image for the camera perspective.
However, as a gaze-contingent method, the compromised quality would be unnoticeable in periphery. Thus, we further refine the visual quality in foveal vision, as detailed in \Cref{sec:method:refine}.
Since we aim to render a novel view in real time and with perceptually high quality, we estimate the foveated image (10 degree), middle peripheral image (60 degree), and full peripheral image (110 degree) from the same novel view for further synthesis. Regarding foveated rendering, we estimate an image with size of 64 x 64, $N_{layer}=12$, $N_{channel}=64$, and $N_{sample}=16$. Regarding middle peripheral and full peripheral rendering, we estimate the two images with size of 256 x 256, $N_{layer}=8$, $N_{channel}=64$, and $N_{sample}=4$. We next upsample middle peripheral to ??? x ??? and full peripheral to 1440 x 1440 for blending.
\nothing{
\subsection{Refinement for Fovea Layer}
\label{sec:method:refine}
Due to the depth bounding and prediction errors, it is particularly challenging to preserve pixel-wise quality without compromising real-time performance. However, the human foveal vision is remarkably sensitive to subtle details across all frequency ranges. To 
To maximize the quality of foveated image, we utilize the limited but consistent network bandwidth by retrieving a local peripheral image from the cloud. The retrieval was realized by cloud-based rendering using users' upstreamed head and gaze parameters.
Due to the inevitable dual-way transmission, it is impossible to instantly obtain the image without latency. Displaying outdated frames have been studied as a major cause of simulator sickness.
}%nothing
\note{describe when to retrieve new full peripheral image.} We leverage local guidance views to fine tune the output foveated image. Local guidance view is decided according to current camera view. We chose the nearest 4 views (or 8 views) as reference views and estimate the color as well as volume density for all reference views. Every time we receive a new peripheral image from the cloud, we calculate the delta between the ground truth and our result for reference views for preparation. Then we warp the delta into target camera view, calculate the average and apply to the fovea result for each frame.

\subsection{Real-time Rendering}
% \subsection{Fovea-Peripheral Blending}
\label{sec:method:blending}
Lastly, we apply alpha blend on the fine-tuned fovea view and middle peripheral view, as well as on the middle peripheral view and full peripheral view via function smo/:

\subsection{Latency-quality joint optimization}
\label{sec:method:optimization}
Increasing rendering sampling naturally improves the image output quality. However, it increases the latency, causing quality drop stretched along time, and more critically, simulator sickness.
Inspired by \cite{Li:2020:TSP}, we perform a spatial-temporal joint optimization to determine the sampling that adapts to individual computational resources. This is achieved via a Pareto efficiency optimization.

$E_{q}\triangleq $

$E_{l}$ is defined as the loss introduced by latency. 
\begin{align}
    \int E_{l} \mathbf{d} t
\end{align}

\begin{figure*}
    \centering
    \subfloat[quality-only]{\includegraphics[width=0.32\linewidth]{example-image-a}}
    \subfloat[latency-only]{\includegraphics[width=0.32\linewidth]{example-image-a}}
    \subfloat[ours]{\includegraphics[width=0.32\linewidth]{example-image-a}}
    \caption{Latency-quality joint-optimization.}
    \label{fig:optimization}
\end{figure*}

%\subsection{Cloud-Based System}
%\label{sec:method:system}
%Finally, we deploy our gaze-contingent synthesizer to a real-time cloud-based streaming system.

>>>>

\section{Method}
\label{sec:method}
In this section, we present our perception-guided neural scene synthesis for virtual reality. Based on a concentric spherical representation (\Cref{sec:method:representation}), our system predominantly comprises of two main steps: visual- and stereo- acuity adaptive view synthesis (\Cref{sec:synthesis}), and image-based real-time rendering to composite stereo display frames (\Cref{sec:method:blending}).
We then craft an analytical spatial-temporal model to guide the optimization of our intra-system variables for desired precision-performance balance (\Cref{sec:method:optimization}).
\input{TOG/fig_notations}
\subsection{Egocentric Neural Representation}
\label{sec:method:representation}
The recent single-object-oriented ``outside-in'' view synthesis methods \cite{sitzmann2019deepvoxels,mildenhall2020nerf} typically represent the training targets using uniform voxelization. However, immersive VR environments introduce open challenges for such parameterization due to the rapid variation of the egocentric (first-person) viewing (e.g., \Cref{fig:teaser:scene}).
As a consequence, the neural representation on large virtual environments typically suffer from ghosting artifacts, low resolution, or slow speed (e.g., the right triangle in \Cref{fig:teaser:quality}).
%We tailor our method specifically for egocentric (first-person) views in immersive VR environments.
%rendering of immersive viewer-centered ``inside-out'' scenes is challenging owing to the rapid variation of scene content. 
\paragraph{Egocentric coordinates}
To tackle this problem we are inspired by the recent panorama dis-occlusion methods \cite{Lin:DeepPanorama,Benjamin:2020:RTV,Broxton:immersiveLF} to depict the rapidly varying first-person views with concentric spheres. This representation enables robust rendering at real-time rates and allows for 6DoF interaction to navigate complex immersive environments. 
As visualized in \Cref{fig:notations}, our representation is parameterized with the number of concentric spheres ($\sphereNum$) and their respective radii ($\mathbf{\sphereRadius}=\{\sphereRadius_i\},i\in [1,\sphereNum]$). Our network aims at predicting the function $\mlpFunc(\hat{\SpatialPt})\triangleq(r,g,b,d)$ of any position $\mathbf{\hat{\SpatialPt}}({\sphereRadius_{i,i\in[1,\sphereNum]},\theta,\phi})$ on these $\sphereNum$ spherical surfaces. %\dnc{$\mathbf{q}$ may be more proper}\qisun{Just to avoid confusion with \Cref{fig:notations}. What do you think?}\dnc{Then perhaps we can use $\hat{q}$ as shown in \Cref{fig:notations} I think the alphabet 'q' means it's in spherical system, 'p' never exists anywhere else}
Here, $(r,g,b)$ and $d$ are the predicted color and density, respectively. 
We then employ a ray marching through this intermediate function to produce images of specified views.

\paragraph{Training and synthesis}
%\dnc{Perhaps we can merge this paragraph with above section? And move the last paragraph of above section to here as a intro. Then this section is more focusing on GAZE-AWARE}
For each scene, we approximate the function $\mlpFunc$ by training a multilayer perceptron (MLP) network, similar to \cite{mildenhall2020nerf}.
The network is composed of $\mlpLayerNum$ fully-connected layers of $\mlpChannelNum$ channels, each using ReLu as the activation function. We will further determine $\mlpLayerNum$, $\mlpChannelNum$ and $\sphereNum$ via a spatial-temporal optimization in \Cref{sec:method:optimization}.
Once trained, the network achieves 4D prediction on the spherical voxels, on which a ray marching rendering is performed to synthesize the final frames for display.% based on users' frame-wise views. 
Similar to \cite{mildenhall2020nerf}, we apply a trigonometric input encoding and compute the L2-distance between the volume-rendered and mesh-rendered images as the training loss function. We discuss the specifics of training data creation and implementations in \Cref{sec:impl}.

\subsection{Gaze-Aware Acceleration}
\label{sec:synthesis}
\qisun{@nianchen: does this section title sounds reasonable to you?}
%The scene representation above is created to parameterize a given 3D space for neural synthesis. 
%the foveated rendering manner of VR display through training two orthogonal neural networks that are tailored for the perceptual acuity variances.
Our concentric spherical coordinate system, as described in \Cref{sec:method:representation}, addresses the view variance problem in VR. However, it sill suffers from significant rendering latency (seconds per frame). This is one of the core challenges causing neural representation to be unsuitable yet for VR. To combat the common performance problem of view synthesis, we predict multiple elemental images based on the viewer's real-time head and gaze motions\nothing{ instead of synthesizing a globally uniform image}. The elemental images are tailored for the visual acuity variances. We detail our network design and elemental image synthesis in the following section. We enable real-time responsiveness via introducing gaze-adaptive synthesis mechanisms.
\input{fig_system}
\paragraph{Adaptive visual acuity} 
%the user's current view $[\rayo,\camDir]$ and gaze position $\gazeDir$ that are tracked in real-time by a VR headset.
Existing neural scene representation networks predict uniform resolution images for offline third-person viewing.
Due to the trade-off between quality and latency during inference, they are typically executed with low quality and/or high latency when it comes large field-of-view and high resolution imagery. \praneeth{this line is confusing} 
Our method, without compromising performance, generates {\it perceptually} high quality frames matching the retinal acuity, i.e. how the retina perceives dynamic and stereo frames.
% Note that foveal/peripheral retinal images requires high/low pixel-wise precision but low/high FoV (number of pixels). Thus, we train two orthogonal neural networks that are tailored for the perceptual acuity variances. 
Because of foveation of imagery by the human visual system \cite{Guenter:2012:F3G}, we train two orthogonal networks (i.e., $\mlpFunc$s) to synthesize foveal ($\imageFoveal(\rayo,\camDir,\gazeDir)$, $0-20$deg), and peripheral elemental images (i.e., two $\mlpFunc$s to deploy for ray marching) for each eye.
The periphery includes mid- ($\imageMid(\rayo,\camDir,\gazeDir)$, $0-45$deg) and high- eccentricity visual fields ($\imageFar(\rayo,\camDir$), $0-110$deg), see \Cref{fig:system}.
Note that $\imageFar$ is independent from the gaze direction $\gazeDir$. 
%The images are defined as functions that maps individual camera rays ($\rayo,\rayd$) to an RGB color.
\dnc{Could move to section 4? I think it's implementation related}\qisun{I think the PDD numbers give readers a rough impression of the varied quality across each of the image, right?} Specifically, to incorporate the display hardware properties and perceptual effects, we define $\imageFoveal$/$\imageMid$/$\imageFar$ as of $128^2$/$256^2$/$230\times256$ resolutions respectively.
That is, the $\imageFoveal$ has the highest spatial resolution of $6.4$ pixels per degree (PDD), higher than those of $\imageMid$ ($5.7$ PDD) and $\imageFar$ ($2.33$ PDD).
%In each frame, the two networks independently return three elemental images, as seen from the separation range in deg, gradually larger areas along eccentricity. An example is shown in \Cref{fig:system}.

\paragraph{Adaptive stereo-acuity}
{
To be rendered on head-mounted displays, stereo elemental images $\imageFoveal^{\{l,r\}}$, $\imageMid^{\{l,r\}}$, and $\imageFar^{\{l,r\}}$ ($l$ and $r$ represent left and right eyes respectively) are required. This doubles the computational cost that is critical for latency- and frame-rate- sensitive VR experience  
\nothing{The separation of the three elemental images considers the varied visual acuity. However, in a head-mounted stereo VR displays, the rendered image are for two eyes, resulting in double computation for $\imageFoveal^{\{l,r\}}$, $\imageMid^{\{l,r\}}$, and $\imageFar^{\{l,r\}}$. Here $l$ and $r$ represent the projection images for left ($\rayo^l$) and right ($\rayo^r$) eyes respectively. 
Whereas, because $\imageMid^{\{l,r\}}$ and $\imageFar^{\{l,r\}}$ demand high spatial resolution due to their large eccentricity coverage, }
}
(please refer to \Cref{sec:study:intra} for detailed comparisons).

To tackle this problem towards real-time experience, we accommodate the inference process with the adaptive stereo-acuity in the periphery.
Specifically, besides the visual acuity, psychophysical studies have revealed human's significantly declined stereopsis while receding from the gaze point \cite{mochizuki2012magnitude}. Inspired this characteristic, instead of inferring $6$ elemental images, we perform the computation with $\imageFoveal^{\{l,r\}}$, $\imageMid^m$, and $\imageFar^m$, where $m$ indicates the view at the midpoint of the left and right eyes. The individual images are further blended for display, as detailed in \Cref{sec:method:blending}. 
\Cref{fig:mono} visualizes the stereo changes from the adaptation.
\input{TOG/fig_mono_comparison}

\subsection{Real-Time Frame Composition}\dnc{May be moved to section 4 (merge with the paragraph 'Integration'). I think it's not closely related to our methodology and we have no key contributions about that. It's more like an engineering issue}\qisun{This whole section? I was concerned that would break the method completeness. Intentionally kept it short according to contribution level here.}
\label{sec:method:blending}
In real-time, we perform a image-based rendering that blends the elemental images. The output frames are rendered on the stereo VR displays.
To accommodate the uniformized $\imageMid^m$, we shift it with $-/+ \frac{\rayo^r-\rayo^l}{2}$ for left/right eye respectively. 
\dnc{Not accurate. shifting is related to mean of fovea depthmap inferred in actual implementation to minimize the misalignment between fovea and mid}
{
With the obtained elemental images as input, a fragment shader is executed to generate final images for each eye. Two adjunct layers are blended using a smoothstep function across $40\%$ of inner layer.
\nothing{After obtaining the individual images, our fragment shader retrieves the three frames and applied a pixel wise bilinear interpolation at each $10\%$ connecting areas between two adjunct layers.}
}
This enhances visual consistency on the edges between layers \cite{Guenter:2012:F3G}.
Moreover, to further preserve peripheral elemental images' visual fidelity due to its low PDD, we enhance the contrast following the mechanism of \cite{Patney:2016:TFR}, the details are visualized in \Cref{fig:method:blending}. %\dnc{Note: the contrast enhancement is performed on elemental images, before blending. Different enhancement parameters are applied for fovea, mid and periph layers}

\begin{figure*}[htb]
    \centering
    \subfloat[network overlay]{\includegraphics[width = 0.3\linewidth]{TOG/figs/system/overlap2.pdf}\label{fig:method:blending:overlay}}\hspace{1em}
    \subfloat[blending]{\includegraphics[width = 0.3\linewidth]{TOG/figs/system/blended2.pdf}\label{fig:method:blending:blending}}\hspace{1em}
    \subfloat[contrast enhance]{\includegraphics[width = 0.3\linewidth]{TOG/figs/system/blended_enhanced2.pdf}\label{fig:method:blending:contrast}}
    \Caption{Our multi-layer real-time rendering system.}
    {%
    \subref{fig:method:blending:overlay} shows the original 3-layer image output overlaid.
    \subref{fig:method:blending:blending} shows the result with cross-layer blending that enables smooth appearance on the edges.
    \subref{fig:method:blending:contrast} shows the contrast-enhanced rendering pass in the periphery.
    }
    \label{fig:method:blending}
\end{figure*}

\subsection{Latency-Quality Joint Optimization}
\label{sec:method:optimization}
As a view synthesis system based on sparse egocentric representation (the density of $\sphereNum$ spheres) and neural networks (the $\mlpLayerNum\times\mlpChannelNum$), the method inevitably introduces approximation errors. On the other hand, these variables also determine the online computational time that involves inferring function $\mlpFunc$ and ray marching the $\sphereNum$ spheres. 
However, VR strictly demands both quality and performance. We present a spatial-temporal model that analytically depicts the correlations and optimizes the variables for ideal user experience.

\paragraph{Precision loss of a 3D scene}
As shown in \Cref{fig:notations}, under the egocentric representation, a 3D point $\pt$ is re-projected as the nearest point on a sphere that connects it to the origin point:
\begin{equation}\label{eq:closestPoint3D}
{\pt^\prime}(\sphereNum,\mathbf{\sphereRadius},\SpatialPt) \triangleq \sphereRadius_k \frac{\pt}{\norm{\pt}}, \ k=\argmin_{j\in[1,\sphereNum]}\left(\norm{\norm{\SpatialPt}-\sphereRadius_j}\right).
\end{equation}

Similar to volume-based representation, the multi-spherical system is also defined in discrete domain. The sparsity thus naturally introduces approximation error that compromises the synthesis quality. To analytically model the precision loss, we investigate the geometric relationship among the camera, the scene, and the representation.
As illustrated in \Cref{fig:teaser:scene,fig:notations},  for a sphere (located at origin point) with radius $\sphereRadius$, its intersection (if exists) with a directional ray $\{\rayo,\rayd\}$ is
\begin{equation}
\intersectionFunc(\sphereRadius,\rayo,\rayd) = \rayo + \left({\left((\rayo\cdot\rayd)^2-\norm{\rayo}^2+\sphereRadius^2\right)}^{\frac{1}{2}}-\rayo\cdot\rayd\right)\rayd,
\label{eq:raySphereIntersection}
\end{equation}
where $\rayo$ and $\rayd$ are the ray's origin point and normalized direction respectively.
% \begin{equation}
% \left\{\intersectionFunc(\sphereRadius_i,\rayo,\rayd) \right\}\ |\ i\in[1,2,\dots,\sphereNum].
% \end{equation}
Inversely, given a view point $\rayo$ observing a spatial location $\SpatialPt$ represented in spherical coordinates, the ray connect them is with the direction $\rayd(\SpatialPt,\rayo)=\frac{\SpatialPt-\rayo}{\norm{\SpatialPt-\rayo}}$. This ray may intersect with more than one spheres. Among them, the closest one to $\SpatialPt$ is:
\begin{equation}\label{eq:closestPointImg}
\begin{aligned}
    \hat{\pt}&(\rayo,\sphereNum, \mathbf{\sphereRadius},\SpatialPt) \triangleq \\
    &\intersectionFunc(\sphereRadius_k,\rayo,\rayd(\SpatialPt,\rayo))\ |\ k=\argmin_{j\in[1,\sphereNum]}\left(\norm{\norm{\SpatialPt}-\sphereRadius_j}\right).
\end{aligned}
\end{equation}
In the 3D space, the offset distance $\norm{\pt^\prime-\hat{\pt}}$ indicates the precision loss at $\pt$ from the representation. By integrating over all view and scene points, we obtain:
\begin{equation}
\begin{aligned}
\sparseError(\sphereNum, \mathbf{\sphereRadius})  &= \iint \norm{\pt^\prime(\sphereNum,\mathbf{\sphereRadius},\SpatialPt)-\hat{\pt}(\rayo,\sphereNum, \mathbf{\sphereRadius},\SpatialPt)} \mathbf{d}\SpatialPt \mathbf{d}\rayo,\\
&\forall \{\rayo, \SpatialPt\} \text{ pair without occlusion in between}.
\label{eq:sparseError}
\end{aligned}
\end{equation}
$\sparseError$ depicts how the generic representation precision of a scene, given a coordinate system defined by $\sphereNum$ and $\mathbf{\sphereRadius}$.

In comparison, a neural representation aims at predicting projected image given a camera position $\rayo$ and orientation $\camDir$. Thus, we further extend \Cref{eq:sparseError} to image space to analyze the error given a set of camera's projection matrix  $\projectionMatrix(\rayo,\camDir)$ as
%However, typical 3D representations are not uniformly distributed points. Moreover, the network only predicts pixel colors/intensities but not depth. By extending \Cref{eq:sparseError} to the captured image space data, we obtain
\begin{align}
\imgSpaceError(\sphereNum, \mathbf{\sphereRadius}, \rayo, \camDir)  = \int \norm{\projectionMatrix(\rayo,\camDir)\cdot\left(\SpatialPt-
\hat{\pt}(\rayo,\sphereNum, \mathbf{\sphereRadius},\SpatialPt)\right)}\mathbf{d}\SpatialPt.
\label{eq:imageError}
\end{align}
From \Cref{fig:notations,eq:imageError}, we observe:
given a fixed min/max range of $\mathbf{\sphereRadius}$, $\sphereNum$ is negatively correlated to $\imgSpaceError$;
with a fixed $\sphereNum$, the correlation between distribution of $\sphereRadius_j$ and scene content (i.e., distribution of $\SpatialPt$s) also determines $\imgSpaceError$.

However, for neural scene representation, infinitely enlarging $\sphereNum$ would significantly raise the challenges in training precision and inference performance.
Likewise, increasing prediction and rendering densities (i.e., higher $\sphereNum$ or $\mlpLayerNum$/$\mlpChannelNum$) naturally improves the image output quality for lower \Cref{eq:imageError} while significantly increases the computation during ray marching, causing quality drop stretched along time. In the performance-sensitive VR scenario, the latency breaks the continuous viewing experiment and may cause simulator sickness. Thus, we determine the coordinate variables with a content-aware optimization towards optimal quality-speed performance.

\paragraph{Spatial-temporal modeling}
Inspired by \cite{Li:2020:TSP,albert2017latency}, we perform a spatial-temporal joint modeling to determine the optimal coordinate system ($\sphereNum$ and $\mathbf{\sphereRadius}$) for positional precision and MLP settings ($\mlpLayerNum$/$\mlpChannelNum$) for color precision that adapt to individual computational resources and scene content. This is achieved via a latency-precision modeling in the spatial-temporal domain.
%Specifically, we model the prediction image $P$ with gaze position ($G$) at time $t$ as $P_G(t)$ and the ground truth retinal image (from a simulated local rendering) as $\hat{P}_G(t)$.
%In practice, increasing the number of spheres $\sphereNum$ or number of MLP layers may affect the computational and transmissional latency but will improve the quality of individual frames (as in \Cref{eq:imageError}). Thus, we define the perceptual error as:
\begin{equation}
\begin{aligned}
\finalError(\sphereNum, &\mlpLayerNum, \mlpChannelNum) = 
\sum_t\int \mlpFunc_{\mlpLayerNum, \mlpChannelNum}(\SpatialPt)\times\\ &\norm{\projectionMatrix(\rayo_t,\camDir_t)\cdot\SpatialPt-\projectionMatrix(\rayo_{t-\latency},\camDir_{t-\latency})\cdot\pt(\sphereRadius_k,\rayo_{t-\latency},\rayd_{t-\latency})}\mathbf{d}\SpatialPt,
%&\int  \norm{ \imgSpaceError(\sphereNum, \mathbf{\sphereRadius}, \rayo_t, \rayd_t) - \imgSpaceError(\sphereNum, \mathbf{\sphereRadius}, \rayo_{t-l(\sphereNum, \mathbf{\sphereRadius})}, \rayd_{t-l(\sphereNum, \mathbf{\sphereRadius})})) } \mathbf{d} t
\label{eq:error:image}
\end{aligned}
\end{equation}
where $\latency\triangleq\latency(\sphereNum, \mlpLayerNum, \mlpChannelNum)$ is the system latency with a given coordinate and network setting.
$\mlpFunc_{\mlpLayerNum, \mlpChannelNum}(\SpatialPt)$ is the L1-distance between the four channels ($r$,$g$,$b$,$a$) and0 the highest values of network layer settings. For simplicity, we assumed uniformly distributed $\mathbf{\sphereRadius}$ with a fixed range of the spherical coverage.

Suggested by Albert et al. \shortcite{albert2017latency}, the latency for a foveated system shall reach below \textasciitilde$50$ms for undetectable artifacts. Given our test device's eye-tracking latency \textasciitilde$12ms$ and the photon submission latency \textasciitilde$14ms$ (\cite{albert2017latency}), the synthesis and rendering latency shall be less than $L_0 = 24$ms.
Thus, we determine the optimal $\{\sphereNum, \mathbf{\sphereRadius}\}$ to balance latency and precision as
\begin{equation}
\begin{aligned}
    &\argmin_{\sphereNum, \mlpLayerNum, \mlpChannelNum} \finalError(\sphereNum, \mlpLayerNum, \mlpChannelNum), \\
    & s.t.\ {l(\sphereNum, \mathbf{\sphereRadius})} < L_0
\end{aligned}
\end{equation}
\Cref{fig:optimization} visualizes an example of the optimization mechanism for in fovea. The optimized results ($\sphereNum, \mlpLayerNum, \mlpChannelNum$) for individual networks are detailed in the next section. The visual quality is validated by psychophysical study (\Cref{sec:study:user}) and objective analysis (\Cref{sec:study:quality}). The latency breakdown of our system is reported in \Cref{sec:study:intra}.
\input{TOG/fig_optimization}
\input{TOG/4.5-implementation}
